An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2017-02-07 icml2017/paper.tex
Updated Appendix B.1
Got the table re-sorted and a description there.

\textbf{Daniel: need to revise this to take into account reviewer suggestion,
with future work ideas. One reviewer suggested
using~\cite{DBLP:conf/icml/ChenG16}.}

We have derived an M-H test for minibatch MCMC which approximates full data
tests. We present theoretical results and experimentally show the benefits of
our test on Gaussian mixtures and logistic regression. Directions for future
work include running more experiments with a particular focus on controlling
variances, testing on neural networks, combining our results with Hamiltonian
Monte Carlo methods, providing a recipe for how to use our algorithm (following
the framework of~\cite{sgmcmc_2015}), or integrating parallel
MCMC~\cite{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts.

@@

% Daniel: moved to appendix so we can expand it. Originally we had only two
% lines, and it looked pretty awkward. So let's add more. Note: this also
% depends on whether we use [-10,10] for the X_c range, or [-20,20], etc...
\begin{table}[t]
\caption{Errors ($L_\infty$) in $X_{\rm norm}+X_{\rm corr}$ versus $X_{\rm
log}$, with $N=4000$ (top row) and $N=2000$ (bottom row).}
\label{tab:xcorr}
\vskip 0.15in
\begin{center}
\begin{small}
\begin{sc}
%% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=4000$ & $\sigma=1.1$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & \textbf{4.3e-3}  \\
10     & 6.0e-3  \\
1      & 1.0e-2  \\
0.1    & 1.2e-2  \\
\belowspace
0.01   & 3.5e-2  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=4000$ & $\sigma=1.0$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 1.9e-3  \\
10     & \textbf{8.9e-4}  \\
1      & 1.6e-3  \\
0.1    & 2.8e-3  \\
\belowspace
0.01   & 5.2e-3  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=4000$ & $\sigma=0.9$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 1.2e-3  \\
10     & 2.6e-4  \\
1      & \textbf{1.0e-4}  \\
0.1    & 2.0e-4  \\
\belowspace
0.01   & 3.9e-4  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=4000$ & $\sigma=0.8$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 8.3e-4  \\
10     & 1.3e-4  \\
1      & 2.5e-5  \\
0.1    & \textbf{6.7e-6}  \\
\belowspace
0.01   & 7.4e-6  \\
\hline
\end{tabular}
%% ---- NEW ROW ----
\vskip 0.2in %% OH this works!
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=2000$ & $\sigma=1.1$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 6.8e-3  \\
10     & \textbf{4.6e-3}  \\
1      & 7.5e-3  \\
0.1    & 1.3e-2  \\
\belowspace
0.01   & 2.4e-2  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=2000$ & $\sigma=1.0$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 4.4e-3  \\
10     & 1.3e-3  \\
1      & \textbf{1.1e-3}  \\
0.1    & 2.0e-3  \\
\belowspace
0.01   & 3.6e-3  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=2000$ & $\sigma=0.9$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 3.3e-3  \\
10     & 6.4e-4  \\
1      & 1.6e-4  \\
0.1    & \textbf{1.3e-4}  \\
\belowspace
0.01   & 2.7e-4  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=2000$ & $\sigma=0.8$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 2.6e-3  \\
10     & 4.0e-4  \\
1      & 6.7e-5  \\
0.1    & 1.4e-5  \\
\belowspace
0.01   & \textbf{5.0e-6}  \\
\hline
\end{tabular}
\end{sc}
\end{small}
\end{center}
\vskip -0.1in
\end{table}

\textbf{Daniel: TODO, figure out what values to test, put optimal values in bold
face here, etc., note that these are with the range $[-20,20]$ except for the
best value in $[-10,10]$; maybe we should decrease the range?}

>>>>

\textbf{Daniel: double check this in case we add a third experiment.}

We have derived an M-H test for minibatch MCMC which approximates full data
tests. We present theoretical results and experimentally show the benefits of
our test on Gaussian mixtures and logistic regression.  Directions for future
work include integrating our algorithm with~\cite{cutting_mh_2014} by applying
both tests each iteration or utilizing the variance reduction techniques
suggested in~\cite{DBLP:conf/icml/ChenG16}.  More elaborate extensions involve
running more experiments on other models such as neural networks and providing a
recipe for how to use our algorithm (following the framework
of~\cite{sgmcmc_2015}).

@@

% Daniel: moved to appendix so we can expand it. Originally we had only two
% lines, and it looked pretty awkward. So let's add more. Note: this also
% depends on whether we use [-10,10] for the X_c range, or [-20,20], etc...
\begin{table}[t]
\caption{Errors ($L_\infty$) in $X_{\rm norm}+X_{\rm corr}$ versus $X_{\rm
log}$, with $N=4000$ (top row) and $N=2000$ (bottom row).}
\label{tab:xcorr}
\vskip 0.15in
\begin{center}
\begin{small}
\begin{sc}
%% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=2000$ & $\sigma=0.8$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 2.6e-3  \\
10     & 4.0e-4  \\
1      & 6.7e-5  \\
0.1    & 1.4e-5  \\
\belowspace
0.01   & \textbf{5.0e-6}  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=2000$ & $\sigma=0.9$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 3.3e-3  \\
10     & 6.4e-4  \\
1      & 1.6e-4  \\
0.1    & \textbf{1.3e-4}  \\
\belowspace
0.01   & 2.7e-4  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=2000$ & $\sigma=1.0$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 4.4e-3  \\
10     & 1.3e-3  \\
1      & \textbf{1.1e-3}  \\
0.1    & 2.0e-3  \\
\belowspace
0.01   & 3.6e-3  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=2000$ & $\sigma=1.1$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 6.8e-3  \\
10     & \textbf{4.6e-3}  \\
1      & 7.5e-3  \\
0.1    & 1.3e-2  \\
\belowspace
0.01   & 2.4e-2  \\
\hline
\end{tabular}
%% ---- NEW ROW ----
\vskip 0.2in %% OH this works!
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=4000$ & $\sigma=0.8$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 8.3e-4  \\
10     & 1.3e-4  \\
1      & 2.5e-5  \\
0.1    & \textbf{6.7e-6}  \\
\belowspace
0.01   & 7.4e-6  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=4000$ & $\sigma=0.9$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 1.2e-3  \\
10     & 2.6e-4  \\
1      & \textbf{1.0e-4}  \\
0.1    & 2.0e-4  \\
\belowspace
0.01   & 3.9e-4  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=4000$ & $\sigma=1.0$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 1.9e-3  \\
10     & \textbf{8.9e-4}  \\
1      & 1.6e-3  \\
0.1    & 2.8e-3  \\
\belowspace
0.01   & 5.2e-3  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=4000$ & $\sigma=1.1$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & \textbf{4.3e-3}  \\
10     & 6.0e-3  \\
1      & 1.0e-2  \\
0.1    & 1.2e-2  \\
\belowspace
0.01   & 3.5e-2  \\
\hline
\end{tabular}
\end{sc}
\end{small}
\end{center}
\vskip -0.1in
\end{table}

In Section~\ref{sec:correction}, we described our derivation of the correction
distribution $C_\sigma$ for random variable $X_{\rm corr}$.
Table~\ref{tab:xcorr} shows our $L_\infty$ error results for the convolution
(Equation~(\ref{eq:overall_corr_problem})) based on various hyperparameter
choices.  We test using $N=2000$ and $N=4000$ points for discretization within
a range of $X_{\rm corr} \in [-20,20]$, covering essentially all the probability
mass. We also vary $\sigma$ from 0.8 to 1.1.

We observe a tradeoff. With smaller $\sigma$, our $C_\sigma$ is closer to the
ideal distribution (as judged by $L_\infty$ error), but this imposes a stricter
upper bound on the sample variance of $\Delta^*$ before our test can be applied,
which thus results in larger minibatch sizes. Conversely, a more liberal upper
bound means we avail ourselves of smaller minibatch sizes, but at the cost of a
less stable derivation for $C_\sigma$.

\textbf{We chose XXXX to use in our experiments, which empirically provided an
acceptable tradeoff to us.} This is reflected in
Algorithm~\ref{alg:our_algorithm}, which assumes that we used \textbf{TODO TODO
TODO TODO TODO} but we reiterate that the choice is arbitrary so long as $0 <
\sigma < \sqrt{\pi^2/3} \approx 1.814$, the standard deviation of the standard
logistic distribution, since there must be some variance left over for $X_{\rm
corr}$.
